---
title: "Case Study - Spotify"
author: "Nishanth Gandhidoss"
date: "October 20, 2016"
output: html_document
---

##Importing the data set

```{r}
df <- read.csv("df.csv")
```

**Removing the noise caused by peak, because when we found the relation between peak and succesfulness of the songs, most of the succesful songs fall under the range of 1 to 10 of peak values and there is only one song which is false in the data. For more convenencie we removed the observation**

```{r}
df1 <- df[!df$X == df[df$Successful == 0 & df$Peak < 10, ]$X, ]
```

##Separating the train data and test data for analysis

```{r}
train_data <- df1[1:300, ]
test_data <- df1[301:nrow(df1), ]
```


##Finding the correlation between the varibles

```{r}
cor(train_data$Successful, train_data$Peak)
cor(train_data$Successful, train_data$danceability)
cor(train_data$Successful, train_data$energy)
cor(train_data$Successful, train_data$key)
cor(train_data$Successful, train_data$loudness)
cor(train_data$Successful, train_data$mode)
cor(train_data$Successful, train_data$speechiness)
cor(train_data$Successful, train_data$acousticness)
cor(train_data$Successful, train_data$instrumentalness)
cor(train_data$Successful, train_data$liveness)
cor(train_data$Successful, train_data$valence)
cor(train_data$Successful, train_data$tempo)
cor(train_data$Successful, train_data$duration_ms)
cor(train_data$Successful, train_data$time_signature)
```

**The correlation coefficient (a value between -1 and +1) tells you how strongly two variables are related to each other.**

##Removing non impact varibles

```{r}
non_impact_cols <- c(1:6, 9, 12, 18:23)
train_data <- train_data[, -non_impact_cols]
test_data <- test_data[, -non_impact_cols]
```

##Scaling the varibles

```{r}
scale_data <- function(data) {
    data.frame(lapply(data, function(data) 
                            scale(data, center = FALSE, 
                                  scale = max(data, na.rm = TRUE)/1)))
}
train_data <- scale_data(train_data)
test_data <- scale_data(test_data)
```

##Building the model

```{r}
model <- glm(Successful ~., family = binomial(link = 'logit'), data = train_data)
summary(model)
```

**We see the deviance residuals, which are a measure of model fit. This part of output shows the distribution of the deviance residuals for individual cases used in the model.**

**For a one unit increase in loudness, the log odds of sucessceful decrease by 0.20.**

```{r}
confint(model)
```

##Checking the fit

```{r}
fit <- predict(model, newdata=test_data, type='response')
fit <- ifelse(fit > 0.5,1,0)
fit
```

##Calculating the accuracy

```{r}
accuracy <- mean(fit != test_data$Successful)
print(paste('Accuracy',1-accuracy))
```

**And thus, we have an accuracy of 61.34%**

